Supervised Learning for Psychological Attention to High School Students

### 2020-11-18 article/article.tex
[ article ] working on methodology

\subsection{Survey}
The used survey for extracting data was initially designed for developing
a Fuzzy Inference System, but the data was labeled according to
unsupervised clustering techniques, obtaining 2 clusters through the subtractive
algorithm \parencite{chiu1994} and then sharpened using Fuzzy-C-Means
\parencite{dunn1973}. The data has 6 variables and 113 observations. This
dataset was divided with the standard $60-20-20$ rule, creating a training
dataset with $60\%$ of the complete data, a validation set with $20\%$ and a
test set with $20\%$ as well. All data was normalized by dividing each dimension
by its maximum (in absolute value), casting the data into the [-1,1] hypercube.

@@

\subsection{Decision Trees}
We refer to \textcite{bramer2007} for the general ideas behind the decision tree
classifier.

@@

The results are presented using three learning curves: the sum of local
gradients per layer at the end of each epoch, the average error energy and a
Receiver Operating Characteristic (ROC) curve, comparing the obtained results
with the real labels.
\begin{itemize}
  \item The sum of gradients per layer at the end of each epoch is an approach
  visualizing if the training algorithm is solving the optimization problem,
  since the sum of local gradients must converge to zero as each gradient
  converges to zero. It is well-known that the vanish of gradients is a
  necessary condition for optimality and the training algorithm is just an
  optimization problem minimizing the output error (cost function).
  \item The average error energy is the cost function of the optimization
  problem. Let $y_i\in\mathbb{R}^n$ be the real output of a given input pattern
  $x_i\in\mathbb{R}^m$ and let $\hat{y}_i\in\mathbb{R}^n$ be the output of the
  NN for the same input pattern. Denoting each component by superscripts, the
  average error energy $\xi_{av}$ on each epoch is then computed as
  \begin{equation}
    \xi_{av}=\dfrac{1}{2}\sum_{i=1}^{N}\sum_{k=1}^n\left(y^k_i-\hat{y}^k_i\right)^2,
  \end{equation}
  where $N$ is the training sample size. Thus, it is clear that
  $\xi_{av}\rightarrow0$ as the training algorithm solves the learning problem
  well.
  \item The ROC curve is a tool that helps visualize the correct classifications
  of the learning machine by plotting the true positive rate on the vertical
  axis, and false positive rate on the horizontal axis. The Area Under the Curve
  (AUC) is an indicator of how good the data is being classified by the model,
  with higher values being better usually better.
\end{itemize}

>>>>

\subsection{Survey}
The used survey for extracting data was initially designed in the developing
a Fuzzy Inference System, but the data was labeled according to
unsupervised clustering techniques, obtaining 2 clusters through the subtractive
algorithm \parencite{chiu1994} and then sharpened using Fuzzy-C-Means
\parencite{dunn1973}. The data has 6 variables and 113 observations. This
dataset was divided with the standard $60-20-20$ rule, creating a training
dataset with $60\%$ of the complete data, a validation set with $20\%$ and a
test set with $20\%$ as well. All data was normalized by dividing each dimension
by its maximum (in absolute value), casting the data into the [-1,1] hypercube.

Remark: the dataset includes both inputs (first three variables) and outputs
(last three variables) of the Fuzzy Inference System. This will be later
useful to organize relationships in the variables for more intuitive plots.

@@

Given that the sample size is fixed, the analysis for the ``Probably
Approximately Correct'' (PAC) learning was not straight-forward, as the
classifier's maximum permited error $\epsilon$ was estimated from the classic
formulae, rather than calculating the adequate sample size for a given $\delta$.

The PAC learning here applied is based on the ideas presented in
\parencite{kubat2017}. Let $N$ be number of samples in the training set,
$\epsilon$ is the classifier maximum permited error rate, $\delta$ is the
probability that a classifier with error rate greater than $\epsilon$ is
\text{error-free} on the training set, $m$ be the dimension of each training
patttern, and let $VC$ be the Vapnik-Chervonenkis dimension of the given
classifier. Hence, for learning in continuous domains, the lower bound for
$N$ is

\begin{equation}
N\geq \max\left\{ \dfrac{4}{\epsilon}\ln\left(\dfrac{2}{\delta}\right),
\dfrac{8\mathrm{VC}}{\epsilon}\ln\left(\dfrac{13}{\epsilon}\right) \right\}.
\end{equation}

It is clear that solving for $\epsilon$ is non-trivial for given values of
$\delta$, $N$ and $\mathrm{VC}$; thus, an optimization approach was used to
estimate $\epsilon$ for the input values, solving the nonlinear optimization
problem

\begin{equation}
\begin{split}
\min_{\epsilon\in\mathbb{R}_+}\ &J=\,\epsilon^2\\
\mathrm{s.t.}\ &\dfrac{4}{\epsilon}\ln\left(\dfrac{2}{\delta} \right)\leq N\\
\quad&\dfrac{8\mathrm{VC}}{\epsilon}\ln\left(\dfrac{13}{\epsilon}\right)\leq N.
\end{split}
\end{equation}

The problem was initially constrained to $\epsilon\in[0,1]$, but it had not
feasible solution for the size of the training set, yielding values of
$\epsilon$ greater than 1. Instead, the constrained was removed and a
sensitivity analysis was made for the training set size, varying the latter to
find a suitable value for $\epsilon$. The size of training set was varied in
$(0.8N, 10N)$ for each kernel SVM and the estimated $\epsilon$ are presented
in Fig. \ref{fig:error-SVM}, with a fixed $\delta$. Note that $\mathrm{VC}=7$
for the linear SVM, since
$m=6$, and $\mathrm{VC}=84$ for the polynomial SVM, since cubic polynomials were
selected; furthermore, the VC dimension for the RBF SVM is infinite and this
case was not included (see \parencite{burges1998}).

Note that $\epsilon$ in the linear SVM enters the feasible value for training
sets above 150 samples and stabilizing around $\epsilon=0.3$ for large training
set samples value. In contrast, $\epsilon$ for the polynomial SVM never reaches
a suitable value below 1, indicating that this classifier requires more than 600
samples in the training set to satisfy the PAC learning.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/svm-linear-error.pdf}
        \caption{Linear SVM.}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/svm-poly-error.pdf}
        \caption{Polynomial SVM}
    \end{subfigure}
    \caption{Estimated $\epsilon$ for different training set sizes.}
    \label{fig:error-SVM}
\end{figure}

\subsection{Decision Trees}
We refer to \textcite{bramer2007} for the general ideas behind the decision tree
classifier. The DTs were used directly from Scikit-Learn
\parencite{scikit-learn, sklearn_api} using
\href{https://bit.ly/3lDHADX}{Julia's API}. The DT were trained using all
default parameters in the suite, but using the ``entropy'' (information gain)
criterion for measuring the quality of splits.

The same idea from the SVM classifiers for plotting the countour of the decision
boundaries of the fitted model was used, plotting the training dataset with the
real labels and the contour for the decision function.

Additionally, it is well-known that the bound for the training dataset size is
given by
\begin{equation}
  N\geq \dfrac{\ln(2)}{2\epsilon^2}\left[(2^k-1)(1+\log_2 m) + 1 + \ln\left(
  \dfrac{1}{\delta}\right)\right],
\end{equation}
but again, the size of the dataset is fixed,

@@

The results are presented using three learning curves: the sum of local
gradients per layer at the end of each epoch, the average error energy and a
Receiver Operating Characteristic (ROC) curve, comparing the obtained results
with the real labels.
\begin{itemize}
  \item The sum of gradients per layer at the end of each epoch is an approach
  visualizing if the training algorithm is solving the optimization problem,
  since the sum of local gradients must converge to zero as each gradient
  converges to zero. It is well-known that the vanish of gradients is a
  necessary condition for optimality and the training algorithm is just an
  optimization problem minimizing the output error (cost function).
  \item The average error energy is the cost function of the optimization
  problem. Let $y_i\in\mathbb{R}^n$ be the real output of a given input pattern
  $x_i\in\mathbb{R}^m$ and let $\hat{y}_i\in\mathbb{R}^n$ be the output of the
  NN for the same input pattern. Denoting each component by superscripts, the
  average error energy $\xi_{av}$ on each epoch is then computed as
\begin{equation}
\xi_{av}=\dfrac{1}{2}\sum_{i=1}^{N}\sum_{k=1}^n\left(y^k_i-\hat{y}^k_i\right)^2,
\end{equation}
  where $N$ is the training sample size. Thus, it is clear that
  $\xi_{av}\rightarrow0$ is a good initial approach for evaluating the
  learning performance.
  \item The ROC curve is a tool that helps visualize the correct classifications
  of the learning machine by plotting the true positive rate on the vertical
  axis, and false positive rate on the horizontal axis. The Area Under the Curve
  (AUC) is an indicator of how good the data is being classified by the model,
  with higher values being better usually better.
\end{itemize}
